### 2024-05-08 output/presentations/update/update.qmd
Started work on hydrology station scraping

---
title: "The negative external effects of deforestation on downstream populations in Brazil"
author: "Felix Schulz"
format: 
  revealjs:
    theme: [default, custom.scss]
    template-partials: 
      - title-slide.html
    slide-number: c/t
title-slide-attributes:
  data-background-image: "./images/Amazonian_Dystopia_site06.jpg"
  data-background-size: cover
width: 1440
height: 960
bibliography: "./Environmental Economics Project.bib"
---

## Project Outline

- [X] Literature Analysis
- [X] Data Acquisition
- [ ] Hydrological Model
- [ ] Statistical Analysis

## Literature

#### Externalities of Deforestation

- Logging has lead to an estimated emission of around 36 gigatonnes of CO2-equivalent greenhouse gas emissions from 2001 to 2022 in Brazil alone [@worldresourcesinstituteGlobalForestWatch2023]
- Research suggest catastrophic biodiversity tipping points[@malhiClimateChangeDeforestation2008]
- Wildfires increase rates of pulmonary diseases [@reddingtonAirQualityHuman2015]
- Malaria incidence appears to be correlated with deforestation [@olsonDeforestationMalariaMancio2010]

#### Soil Erosion, River Pollution and Health Outcomes

- Soil properties are substantially altered after logging [@veldkampDeforestationReforestationImpacts2020]
- Increased erosion and agricultural use may deteriorate drinking water quality downstream [@maHumanHealthRisk2014]
- Studies estimate significant increase in defensive expenditures due to erosion [@clarkErodingSoilsOfffarm1985]

## Data Acquisition

- Obtained data on 
  - Deforestation (annual, 30m-resolution maps)
  - River pollution (daily, >4000 sensor stations)
  - Child mortality (daily, by postcode)
  - Cloud coverage (annual, 90m-resolution maps)

#### Identification

- Monitoring is satellite-based (DETER project) [@nepstadSlowingAmazonDeforestation2014]
- Cloud cover as a potential instrument for deforestation

## Hydrological Model I

- Established a river model

<br>

:::: {.columns}

::: {.column width="45%"}

![Example of River Network](images/river_network_example.png)

:::

::: {.column width="45%"}

![Watersheds of Brazilian Rivers](images/rivers_brazil.png)

:::

::::

## Hydrological Model II

- Computed detailed drainage-dependent query polygons

:::: {.columns}

::: {.column width="45%"}

![Example of Detailed Watershed Computation](images/drainage_example.png)

:::

::: {.column width="45%"}

![Example of Extraction Polygons](images/example_subsplitting.png)

:::

::::

## Statistical Analysis

- Descriptives and experiments with aggregation and weighting
- Exploring the correlation of deforestation and pollution readings at sensors
- Estimating panel- and spatial models with IV with health outcomes (Spatial-Lag/Durbin Model)

## Sources

### References

:::{#refs}
:::

### Images
- Title slide: [Amazonian Dystopia, Lalo de Almeida](https://lalodealmeida.com.br/amazonian_dystopia/)
- Slide X: DALL·E 3